---
title: "Red Wine"
author: "Yoon-gu Hwang"
date: "June 30, 2015"
output: html_document
---

```{r}
library(ggplot2)
```

# Univariate Plot Section
```{r : Load *.csv file.}
setwd('~/repos/nanodgreep3')
wdf <- read.csv('wineQualityReds.csv')
wdf <- wdf[,-1] # Get rid of unused column : 'X'
str(wdf)
summary(wdf)
```

## Quality
```{r}
qplot(wdf$quality) +
  xlab('quality')
```

It is better to convert qulaity variable to categorical variable, because its values are always integers.

```{r}
wdf$quality <- factor(wdf$quality)
qplot(wdf$quality) +  xlab('quality')
```

```{r}
summary(wdf$quality)
```

Quality has 6 scores, from 3 to 6. Most of all have 5 or 6.

## Fixed Acidity
```{r}
summary(wdf$fixed.acidity)
qplot(wdf$fixed.acidity) + xlab('Fixed Acidity')
```

## Volatile Acidity
```{r}
summary(wdf$volatile.acidity)
qplot(wdf$volatile.acidity) + xlab('Volatile Acidity')
```

```{r}
summary(wdf$volatile.acidity)
qplot(wdf$volatile.acidity) + xlab('Volatile Acidity') +
  scale_x_continuous(limits=c(0, 1.2))
```

## Citric Acid 
```{r}
summary(wdf$citric.acid)
qplot(wdf$citric.acid) + xlab('Citric Acid')
```

```{r}
summary(wdf$citric.acid)
qplot(wdf$citric.acid) + xlab('Citric Acid') +
  scale_x_continuous(limits=c(0.0, 0.8))
```

## Residual Sugar
```{r}
summary(wdf$residual.sugar)
qplot(wdf$residual.sugar) + xlab('Redidual Sugar')
```

```{r}
summary(wdf$residual.sugar)
qplot(wdf$residual.sugar) + xlab('Redidual Sugar') +
  scale_x_continuous(limits = c(0, 5))
```

## Chlorides
```{r}
summary(wdf$chlorides)
qplot(wdf$chlorides) + xlab('Chlorides')
```

```{r}
summary(wdf$chlorides)
qplot(wdf$chlorides) + xlab('Chlorides') +
  scale_x_continuous(limits = c(0,0.2))
```

## Free Sulfur Dioxide
```{r}
summary(wdf$free.sulfur.dioxide)
qplot(wdf$free.sulfur.dioxide) + xlab('Free Sulfur Dioxide')
```

```{r}
summary(wdf$free.sulfur.dioxide)
qplot(wdf$free.sulfur.dioxide) + xlab('Free Sulfur Dioxide') +
  scale_x_continuous(limits = c(0,60))
```

## Total Sulfur Dioxide
```{r}
summary(wdf$total.sulfur.dioxide)
qplot(wdf$total.sulfur.dioxide) + xlab('Total Sulfur Dioxide')
```

```{r}
summary(wdf$total.sulfur.dioxide)
qplot(wdf$total.sulfur.dioxide) + xlab('Total Sulfur Dioxide') +
  scale_x_continuous(limits = c(0,160))
```

## Density
```{r}
summary(wdf$density)
qplot(wdf$density  ) + xlab('Density')
```

```{r}
summary(wdf$density)
qplot(wdf$density) + xlab('Density') +
  scale_x_continuous(limits = c(0.990,1.005))
```

## pH
```{r}
summary(wdf$pH)
qplot(wdf$pH) + xlab('pH')
```

```{r}
summary(wdf$pH)
qplot(wdf$pH) + xlab('pH') +
  scale_x_continuous(limits = c(2.6,3.8))
```

## Sulphates
```{r}
summary(wdf$sulphates)
qplot(wdf$sulphates) + xlab('Sulphates')
```

```{r}
summary(wdf$sulphates)
qplot(wdf$sulphates) + xlab('Sulphates') +
  scale_x_continuous(limits = c(0.3,1.5))
```

## Alcohol
```{r}
summary(wdf$alcohol)
qplot(wdf$alcohol) + xlab('Alcohol')
```

```{r}
summary(wdf$alcohol)
qplot(wdf$alcohol) + xlab('Alcohol') +
  scale_x_continuous(limits = c(8,14))
```

## Univariate Analysis

# Bivariate Plots Section

```{r}
pairs(~fixed.acidity + volatile.acidity + citric.acid + residual.sugar, 
      data = wdf, main = "Example")
```

