---
title: "Red Wine"
author: "Yoon-gu Hwang"
date: "July 5, 2015"
output: html_document
---

```{r, preparation}
library(ggplot2)
library(GGally)
library(gridExtra)
library(psych)
library(dplyr)
library(scales)
library(memisc)
theme_set(theme_minimal(20))
```

# Univariate Plot Section
```{r : Load *.csv file.}
setwd('~/repos/nanodgreep3')
wdf <- read.csv('wineQualityReds.csv')
wdf <- wdf[,-1] # Get rid of unused column : 'X'
dim(wdf)
names(wdf)
str(wdf)
levels(factor(wdf$quality))
summary(wdf)
```

The average quality is 5.6 and median of quality is 6.0. 
About 75% of wines have quality score under or equal to 6.
About 75% of wines have fixed acidity less than 10.0.
About 75% of wines have residual sugar less than 2.6, but its maximum value, 15.5 is very high(i.e. very sweet).
All wines have similar density, from 0.99 to 1.00.

```{r, Quality}
qplot(x = quality, data = wdf)
```

Quality have all integer values. 
Most wines have 5 or 6 score value.
It can be categorized as three grades.

```{r, Add new variable : taste}

wdf['taste'] <- "GD"
wdf$taste[wdf$quality == 3 | wdf$quality == 4] <- "NG" 
wdf$taste[wdf$quality == 5 | wdf$quality == 6] <- "GD" 
wdf$taste[wdf$quality == 7 | wdf$quality == 8] <- "EX" 
wdf$taste <- factor(wdf$taste, levels=c("NG", "GD", "EX"))
```

- 3 and 4 : Not Good(NG) 
- 5 and 6 : Good(GD)
- 8 and 7 : Excellent(EX)

```{r, taste}
qplot(x = taste, data = wdf)
summary(wdf$taste)
str(wdf$taste)
```

```{r, Percentage taste}
ggplot(wdf, aes(x = taste)) +  
  geom_bar(aes(y = (..count..)/sum(..count..))) + 
  ylab('Percentage')

```

- NG : 4%
- GD : 82%
- EX : 14%

```{r, fixed.acidity}
summary(wdf$fixed.acidity)
qplot(x = fixed.acidity, data = wdf, binwidth=0.25)
```

Most wines have a fixed.acidity between 7 and 14.

```{r, volatile.acidity}
summary(wdf$volatile.acidity)
qplot(x = volatile.acidity, data = wdf)
qplot(x = volatile.acidity, data = wdf, binwidth=0.01) + 
  coord_cartesian(xlim = c(0.1, 1.0))
```

Again, most of wines have a volatile.acidity between 0.2 and 1.0.

```{r, citric.acid}
summary(wdf$citric.acid)
qplot(x = citric.acid, data = wdf)
qplot(x = citric.acid, data = wdf, binwidth=0.01)
table(wdf$citric.acid == 0)
```

About 10% of wines have no citric.acid.

```{r, residual.sugar}
summary(wdf$residual.sugar)
qplot( x= residual.sugar, data = wdf)
qplot( x= residual.sugar, data = wdf, binwidth = 0.1) +
  coord_cartesian(xlim = c(0.9, 12.0))

```

Residual sugar of most wines varies from 1.0 to 7.0.

```{r, chlorides}
summary(wdf$chlorides)
qplot(x=chlorides, data = wdf)
qplot(x=chlorides, data = wdf, binwidth = 0.001) + 
  coord_cartesian(xlim = c(0.05, 0.2))
```

Most wines have a chlorides ranging from 0.05 to 0.2.

```{r, free.sulfur.dioxide}
summary(wdf$free.sulfur.dioxide)
qplot(x=free.sulfur.dioxide, data=wdf)
qplot(x=free.sulfur.dioxide, data=wdf, binwidth=1) +
  coord_cartesian(xlim=c(0, 60))
```

Most wines have free.sulfur.dioxide under 40.

```{r, total.sulfur.dioxide}
summary(wdf$total.sulfur.dioxide)
qplot(x=total.sulfur.dioxide, data=wdf)
qplot(x=total.sulfur.dioxide, data=wdf, binwidth = 1) + 
  coord_cartesian(xlim=c(5, 160))
```

Again, most wine have total.sulfur.dioxide from 5 to 160.

```{r, density}
summary(wdf$density)
qplot(x=density, data=wdf)
qplot(x=density, data=wdf, binwidth=0.0001)
var(wdf$density)
```

Most wine have almost same density, because its variance is 3.5e-06, so small.

```{r, pH}
summary(wdf$pH)
qplot(x=pH, data=wdf)
qplot(x=pH, data=wdf, binwidth = 0.01)
```

Of course, all wines are acidic(i.e. under pH 7), because all wines have pH from 2.5 to 4.010.

```{r, sulphates}
summary(wdf$sulphates)
qplot(x=sulphates, data=wdf)
qplot(x=sulphates, data=wdf, binwidth = 0.01) +
  coord_cartesian(xlim=c(0.3, 1.5))
```

Also, most wines' sulphates have range from 0.3 to 1.2.

### What is the structure of your dataset?

  1. 1599 wines in the data set with 13 features(I added one more feature, 'taste')
    - fixed.acidity
    - volatile.acidity
    - citric.acid
    - residual.sugar
    - chlorides
    - free.sulfur.dioxide
    - total.sulfur.dioxide
    - density
    - pH
    - sulphates
    - alcohol
    - quality
    - taste
  1. Every feature has numerical type except taste. taste have ordered factor with 3 levels, "NG", "GD", and "EX".
  
### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
cor(wdf[, c(1, 5, 6, 7, 8, 9, 10, 12)])
```

```{r}
set.seed(10)
pairs.panels(sample_n(wdf[1:11], 1000), pch=".")
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
